partial: An R Package for Creating Partial Dependence Plots

### 2016-08-11 manuscript/greenwell.tex
minor edits

Setting up a parallel backend is rather straightforward. To illustrate, the following snippet of sets up the \code{partial} function to run in parallel on Unix-like systems\footnote{This example will not run on Windows.} using the \pkg{doParallel}.
\begin{example}
library(doParallel)  # load parallel backend
registerDoParallel(cores = 4)  # use 4 cores
\end{example}
Now, to run \code{partial} in parallel, all we have to do is invoke the \code{.parallel = TRUE} option and the rest is taken care of by the internal call to \pkg{plyr} the parallel backend we loaded. This is illustrated in following snippet of code which obtains the partial dependence of median home value on $\boldsymbol{z}_s = \{$\code{rm}, \code{ptratio}, \code{chas}$\}$ in parallel. The result is displayed in Figure~\ref{fig:partial_par}.
\begin{example}
partial(boston.rf, pred.var = c("rm", "ptratio", "chas"), grid.resolution = 20,
        chull = TRUE, .parallel = TRUE, plot = TRUE)
\end{example}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{partial_par}
  \caption{Partial dependence of \code{cmedv} on \code{lstat}, \code{rm}, and \code{ptratio}. Since \code{ptratio} is contiuous, it is first converted to a shingle; in this case, four groups with with 10\% overlap.}
  \label{fig:partial_par}
\end{figure}

>>>>

To illustrate, we will use the Los Angeles ozone pollution data described in ????. The data are available from \url{http://statweb.stanford.edu/~tibs/ElemStatLearn/datasets/LAozone.data}. The following code chunk loads the data into R:
\begin{example}
ozone <- read.csv(paste0("http://statweb.stanford.edu/~tibs/ElemStatLearn/",
                         "datasets/LAozone.data"), header = TRUE)
\end{example}

Next, we use MARS to to study the relationship between atmospheric ozone (\code{ozone}) concentration and meteorology in the Los Angeles basin in 1976, allowing for up to three-way interactions.
\begin{example}
ozone.earth <- earth(ozone ~ ., data = ozone, degree = 3)
summary(ozone.earth)
\end{example}
A single three-way interaction term between \code{wind}, \code{temp}, and \code{dpg} was selected by the MARS algorithm. To understand this interaction, we can use partial. However, since the partial dependece between three continuous variables can be computationally expensive, we will run \code{partial} in parallel.

Setting up a parallel backend is rather straightforward. To illustrate, the following snippet of sets up the \code{partial} function to run in parallel on Unix-like systems\footnote{This example will not run on Windows.} using the \pkg{doParallel} package.
\begin{example}
library(doParallel)  # load parallel backend
registerDoParallel(cores = 4)  # use 4 cores
\end{example}

Now, to run \code{partial} in parallel, all we have to do is invoke the \code{.parallel = TRUE} option and the rest is taken care of by the internal call to \pkg{plyr} and the parallel backend we loaded. This is illustrated in following snippet of code which obtains the partial dependence of \code{ozone} on \code{wind}, \code{temp}, and \code{dpg} in parallel. The result is displayed in Figure~\ref{fig:partial_par}.
\begin{example}
partial(ozone.earth, pred.var = c("wind", "temp", "dpg"), plot = TRUE,
        chull = TRUE, .parallel = TRUE)
\end{example}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{partial_par}
  \caption{Partial dependence of \code{ozone} on \code{wind}, \code{temp}, and \code{dpg}. Since \code{dpg} is contiuous, it is first converted to a shingle; in this case, four groups with 10\% overlap.}
  \label{fig:partial_par}
\end{figure}

% \begin{itemize}
%   \item \code{wind}: wind speed (mph) at Los Angeles International Airport (LAX)
%   \item \code{temp}: temperature (degrees F) at Sandburg Air Force Base
%   \item \code{dpg}: the pressure gradient (mm Hg) from LAX to Dagget, CA
% \end{itemize}

@@
